\section{Discussion}

\subsection{Why the Stage-3 surrogate is the most credible screening model}

Stage~1 is necessary because it provides a stable mean trajectory on representative curves and validates the $A$-scaling hypothesis before the full dataset is introduced. On its own, however, it ignores most within-condition variation and therefore cannot be the final screening model.

Stage~2 is the first model that learns a conditional distribution on the full filtered dataset. Its role is to separate conditional mean from conditional spread and to provide the teacher distribution used for sparse late-time observations. This is why its MAE is slightly larger than that of Stage~1: the supervision problem is harder and more realistic.

Stage~3 is the most credible screening model because it combines the strengths of both earlier stages. Where raw observations are dense, it learns directly from measured CDF data; where observations become sparse, it falls back on the Stage~2 teacher, following the right-censoring rationale established in Chapter~2. That hybrid supervision explains both the improved early-time fidelity and the more conservative late-time behaviour seen in the final results.

\subsection{Validated findings and scope limitations}

Within the declared scope, four claims are supported by the results. First, the optical workflow can extract repeatable penetration-oriented targets from multihole Mie videos at archive scale. Second, the reduced-order fitting stage provides both smoothing and practical target construction for censored observations. Third, the staged surrogate improves from representative mean prediction to uncertainty-aware full-dataset prediction and finally to raw-CDF-refined screening performance. Fourth, the final uncertainty intervals are conservative enough to support cautious early-stage ranking rather than only point prediction.

The scope of the claim is intentionally limited.
The extracted spray boundary is interpreted as a Mie-defined liquid-envelope observable under a specific optical protocol,
not as full in-cylinder liquid truth \cite{ecn_liquid_penetration,ecn_liquid_penetration_accessory}.
The model is intended mainly for interpolation within the supported operating domain, not for unconstrained extrapolation.
It does not replace CFD, engine testing, or validated spray--wall impingement physics.
Its role is narrower and more practical: to provide an auditable early-stage engineering screening layer that narrows the design space before higher-fidelity validation begins.

An additional limit lies in the target definition itself. The reduced-order fit converges toward a quarter-power-dominated continuation, and the neural surrogate is regularised to remain monotone increasing. These choices are deliberate shape priors: they enforce intuitively plausible trajectory growth and help repair short-horizon extrapolation near the end of the observation window. They do not, however, model the signal-formation physics of Mie scattering. When the optical signature weakens before the plume reaches the image boundary, the fitted target can compensate by continuing the penetration trajectory. The resulting penetration should therefore be interpreted as a continuation of a Mie-defined geometric envelope rather than as direct evidence of a strong liquid-phase tip. It does not directly quantify local liquid mass, droplet size, phase state, or wall-film potential \cite{ecn_liquid_penetration,ecn_liquid_penetration_accessory}.

This extrapolation assumption currently lacks direct control validation. The originally planned validation path was to use wide-field experimental videos of single-plume sprays under identical nozzle types and similar operating conditions (non-evaporating, non-reacting Mie-imaged diesel), provided by the industrial partner. With a field of view approximately twice as wide as the multihole archive in this work, the entire penetration trajectory up to the plateauing phase could be fully observed. On such data, an artificial spatial truncation could be introduced (e.g., taking half the field width), using only the time series before the truncation point to fit the q1 model, and then comparing the extrapolated curve with the complete measured trajectory after the truncation point. This would quantify the reliable range and failure modes of the extrapolation through point-by-time residuals---including systematic directions of overestimation or underestimation, and typical time spans of failure from the truncation point. This experimental design is a direct empirical test of the core assumption of the KD chain in this thesis: that the parametric form $\sigma\!\left((t-t_0)/s\right)\cdot k_{1/4}\,t^{1/4}$ remains a credible continuation beyond the field of view boundary, rather than a bias dominated by shape priors. Because the industrial partner was unable to provide such single-plume data, the above validation could not be executed within the scope of this study; the reliable range of the $q1$ extrapolation thus remains an open empirical question and should be prioritized when subsequent experimental conditions permit.

This distinction matters for downstream wall screening. Spray--wall interaction is not uniformly harmful: if most of the liquid phase has evaporated before the wall is reached, the remaining wall interaction may be less damaging and can in some regimes even assist entrainment and mixture preparation. The regime of primary concern is direct liquid-wall contact and persistent wall-film formation \cite{MorieiraMotaPanao2010,liu2024marine_impingement,peraza2022swi,ahmad2025fueldilution}. The present model should therefore be read as a conservative screening surrogate for possible near-wall reach of a Mie-defined spray envelope, not as a phase-resolved predictor of harmful liquid-wall impingement.

A further limitation lies in the downstream engine-geometry coupling. In the current prototype, piston motion only translates the wall mask through slider-crank kinematics; it does not feed back into the penetration surrogate through time-varying chamber pressure, pressure-wave dynamics, or large-scale gas motion. This simplification is inherited from the OSCC measurements, where each spray event is recorded in a rigid chamber at fixed back-pressure. The surrogate therefore learns penetration under quasi-static chamber conditions rather than under fully coupled in-cylinder transients. Because diesel-spray penetration is sensitive to ambient-gas density, that decoupling should be understood as a pragmatic modelling choice rather than as evidence that pressure and flow effects are negligible \cite{NaberSiebers1996}.

Within the intended use, however, the approximation remains defensible as a first-order engineering screen. The target use here is a short pilot-injection window on the order of milliseconds near top dead centre, where the current toy piston setup moves by less than \SI{1}{mm} over the reported \SI{5}{ms} horizon, and where the immediate need is geometric ranking rather than crank-angle-resolved in-cylinder truth prediction. For low-speed or moderate-speed large-bore dual-fuel engines, that short-window decoupling can therefore be acceptable for conservative screening. The approximation weakens as engine speed rises, injection timing moves farther from top dead centre, or transient in-cylinder pressure and gas velocity materially alter the ambient state seen by the spray. Under those conditions, the present wall-interaction score should be read only as a decoupled screening proxy, not as a coupled penetration predictor.

A related limitation concerns the treatment of spray cone angle. Although cone-angle proxies are extracted in the image-processing stage, they were not analysed systematically across campaigns and were not promoted to a learned target in the final surrogate. In the downstream wall-interaction model, the half-cone angle $\alpha$ therefore enters as a fixed input parameter rather than as a validated time-varying observable. This is a non-trivial simplification: prior spray studies show that spreading angle can vary with transient spray development and operating condition, with ambient-gas density, injection conditions, and nozzle geometry all contributing to the observed behaviour \cite{NaberSiebers1996,JungManinSkeenPickett2015}. The current geometry screen therefore learns uncertainty only along the axial penetration direction; the transverse envelope governed by $\alpha$ is assumed rather than predicted.

The same results also mark clear limits:
\begin{itemize}
  \item The data are room-temperature, non-reacting, and tied to a specific optical definition of the liquid envelope; they do not validate evaporation, ignition, or real wall-film physics.
  \item The surrogate is trained on one supported operating domain and should not be trusted equally under genuine out-of-distribution extrapolation.
  \item The downstream geometry model uses a rotated Gaussian envelope, simplified piston kinematics, and a fixed half-cone angle, while leaving piston-driven pressure and gas-motion feedback decoupled from penetration prediction; it is a near-wall screening proxy, not a production impingement solver.
  \item Campaign-level variation and the hard HZ\_Nozzle3 example show that nozzle-family-dependent early transient dynamics are still only partially captured.
  \item The training target uses the CDF quantile frontier definition rather than the binarized polar penetration (the polar binary penetration metric). For normal injection events, the polar binary penetration metric shows better cross-hole consistency and tracks closer to the actual spray tip, yielding systematically higher penetration estimates than the CDF method. The current surrogate output should be understood as the effective penetration under the CDF definition, not the absolute maximal liquid front; adopting different measurement definitions is a known open direction for applications sensitive to absolute penetration values.
  \item The heteroscedastic Gaussian assumption treats inter-plume penetration variance as a symmetric distribution, but penetration itself is a strictly positive quantity and is dominated by multiplicative injection pressure perturbations. In the Hiroyasu-Arai framework, the multiplicative uncertainty of $\Delta P^{1/4}$ theoretically drives $S(t)$ towards a log-normal distribution rather than a symmetric Gaussian. Experimentally, the cross-plume distribution of the q1 fitting amplitude $k_{1/4}$ is significantly right-skewed (median approx. \SI{69}{mm}, right tail extending to approx. \SI{140}{mm}), and the full-dataset evaluation residual distribution is also skewed to the negative side. The final surrogate's $1\sigma$ coverage (72.2\%) is slightly higher than the theoretical Gaussian value (68.3\%), indicating that the model compensates for the underestimated right tail by slightly inflating $\sigma$---a typical adaptive behavior of symmetric distribution assumptions when faced with right-skewed true distributions. The engineering implication is that the current impingement probability tends to be systematically underestimated for extreme large-penetration events; for early-stage screening used for relative ranking, this bias direction is on the conservative side.
  \item Metallic chamber hardware and the nozzle body produce occlusion and an inhomogeneous background relative to a clean through-path optical setup. The proportion of plume frames affected and the marginal RMSE contribution of occluded conditions have not been quantified, so the reported aggregate error implicitly averages over both occluded and unoccluded regions of the image plane.
  \item The CDF quantile threshold $q$ used to define the penetration front is set close to one but its exact value enters the target definition without a reported sensitivity sweep. Varying $q$ within plausible bounds shifts the extracted target by an unknown millimetre-scale amount that is not propagated into the surrogate's reported error.
  \item The Hiroyasu--Arai amplitude $A=(\Delta P/\rho_a)^{1/4}d_n^{1/2}$ collapses most diameter and pressure variation but does not eliminate it: the sparse-feature $R^2$ on the non-dimensional target is 0.14 rather than zero. The residual condition-dependent structure is delegated entirely to the MLP and has not been examined separately for nozzle-batch confounds.
  \item The onset side-output is supervised against a linear ramp to $t_{\mathrm{ramp}}=\SI{0.2}{ms}$. Both the ramp shape and the cutoff are engineering choices motivated by the typical pilot-injection rise time; their sensitivity has not been ablated, and the resulting $P_{\mathrm{onset}}(t)$ trace should be read as a coarse timing indicator rather than a calibrated probability.
  \item Hole-to-hole interaction and sac-flow coupling are documented in the literature but are not modelled here: the surrogate emits a per-plume conditional distribution and the downstream collision screen treats plumes as independent rotated Gaussians. Plume independence is therefore a modelling convenience inherited from the per-strip extraction pipeline, not an experimental finding.
  \item Forward Kullback--Leibler divergence is used in the Stage~3 distillation objective for consistency with the Stages~1--2 NLL form. The claim that forward and reverse KL produce near-identical predictions in the operating range that matters is asserted from preliminary inspection rather than from a documented ablation, so the directionality choice should be regarded as deliberate but not formally justified.
  \item Group-aware splits prevent leakage at the operating-condition level. The additional leave-one-nozzle-family evaluation shows that the staged MLP remains below the two physical baselines in RMSE on held-out nozzle families, but the hardest held-out family still shows under-coverage. Thus, cross-family transfer has been partially stress-tested, but not solved as a fully reliable out-of-distribution predictor.
  \item The external-cleaning evaluation reports four repeated anchor-off runs, with a seed-to-seed RMSE standard deviation of \SI{0.120}{mm}. Bootstrap confidence bands are still not provided, so the repeated-run mean is more reliable than a single point estimate but should not be read as a full statistical certification.
  \item The staged surrogate is benchmarked against calibrated Hiroyasu--Arai, delayed Naber--Siebers, and sparse heteroscedastic Gaussian-process baselines on the same external-cleaning protocol, as well as against its own Stage~3 ablations. Lighter deterministic regressors such as random forests or gradient boosting are still not included, so the marginal value of the neural architecture against those alternatives remains unquantified.
  \item Held-out RMSE and MAE are computed over $519{,}570$ pointwise samples drawn from $21{,}470$ trajectories, but errors within a trajectory are temporally correlated. The effective number of independent samples is therefore substantially smaller than the raw point count, and confidence statements derived from pointwise sample sizes should be read with that correlation in mind.
\end{itemize}

\subsection{Future validation pathway for the probabilistic screening prototype}

The screening prototype described in Section~4.7 operationalises the probabilistic proxy introduced in the literature review, but the resulting frame-by-frame collision probability scores $P_{\mathrm{hit}}(t)$ have not yet been compared against physical spray--wall observations. This gap is a well-defined future work item rather than an open-ended limitation, because the required reference data exist within the industrial partner's experimental portfolio.

The natural first validation target is high-speed optical imaging of spray injection in a representative piston geometry under engine-representative boundary conditions. In such experiments, frame-resolved observations of spray spreading near the piston crown are directly comparable to $P_{\mathrm{hit}}(t)$: the first frame at which spray contact is observed provides an empirical bound on the collision onset, while the cumulative spray exposure over the injection window can be compared to the integrated score $E = \int P_{\mathrm{hit}}\,\mathrm{d}t$. This comparison would test the axial-penetration surrogate under moving-geometry conditions that the rigid-chamber training dataset cannot supply.

A complementary validation path uses CFD. High-fidelity spray computations resolve the liquid-phase distribution, wall-film mass, and local droplet concentration that the Mie-defined surrogate approximates with a rotated Gaussian envelope. Rather than replacing CFD, the surrogate is intended as a low-cost upstream screen: operating points that score above a collision-probability threshold advance to full computation, while those scored safely below are filtered out. Validating this tiered workflow --- screening recall rate, false negative rate, and the associated computational savings --- requires a set of labelled CFD reference cases that the present dataset does not yet include.

Both paths would also resolve a structural open question in the current prototype: whether the fixed half-cone angle $\alpha$ is an acceptable simplification or whether a time-varying lateral spread model, informed by the cone-angle proxies extracted in Chapter~3, materially alters the collision-probability output. Sensitivity to $\alpha$ can only be assessed reliably once frame-matched reference data are available.

