\section{Conclusions}

Chapter~1 framed this thesis as a six-stage pipeline carried end to end (from a raw optical archive to a usable design-triage tool) through data extraction, censoring-aware curation, physics-informed target construction, staged uncertainty-aware training, multi-protocol evaluation, and screening deployment. This closing chapter returns to that sequence to state what has been validated at each stage and where the prototype's boundaries lie.

This thesis demonstrates that large-scale archived Mie scattering videos of multi-hole diesel injection can be transformed into a calibrated, uncertainty-aware penetration surrogate. The resulting model supports early-stage geometric screening without requiring computational fluid dynamics (CFD) for every candidate.

The main contribution is the workflow rather than the specific neural network backbone. The workflow combines right-censoring-aware target construction, plume-level heteroscedasticity, physics-informed amplitude $A$ scaling, raw CDF/teacher knowledge distillation, and a probabilistic wall-impingement interface. Under a unified evaluation protocol, the final multi-stage MLP substantially outperforms two physical baselines and gives more conservative in-distribution uncertainty coverage than the SVGP baseline. At the same time, the SVGP is the stronger point-wise regressor in both the in-distribution and leave-one-nozzle-out evaluations.

The MLP backbone is therefore retained for three structural reasons: its onset-CDF output head, its regime-aware distillation chain, and its low computational cost for the ablation evidence in Chapters~4 and~5. The broader multi-stage surrogate should be understood as a regressor-agnostic instance of the right-censoring-aware learning protocol. The same protocol can host kernel mean functions as well; the LONO results above (\SI{10.16}{mm} vs.\ \SI{12.55}{mm} mean RMSE) suggest that a kernel mean function would improve out-of-distribution accuracy within this same protocol.

\subsection{Validated Results}

Table~\ref{tab:conclusions-scorecard} summarizes the final validated surrogate performance on the primary uncensored CDF protocol and the leave-one-nozzle-out (LONO) transfer protocol; the points below give the interpretation.

\begin{table}[h]
\centering
\footnotesize
\setlength{\tabcolsep}{5pt}
\begin{tabular}{lcccc}
\toprule
 & \multicolumn{2}{c}{Three-stage MLP} & \multicolumn{2}{c}{SVGP} \\
\cmidrule(lr){2-3}\cmidrule(lr){4-5}
Evaluation protocol & RMSE [mm] & 1$\sigma$/2$\sigma$ & RMSE [mm] & 1$\sigma$/2$\sigma$ \\
\midrule
In-domain, uncensored CDF      & 4.185 & 0.888/0.990 & 4.087 & 0.717/0.946 \\
LONO, 5-fold mean              & 12.55 & 0.575/0.770 & 10.16 & 0.628/0.821 \\
LONO, 4-fold (excl.\ Nozzle0)  & 7.00  & 0.696/0.923 & 5.95  & 0.697/0.890 \\
\bottomrule
\end{tabular}
\caption{Final validated surrogate scorecard on the conservative uncensored CDF protocol (540{,}367 points; 5-seed means, seed std $\le\SI{0.04}{mm}$ RMSE) and leave-one-nozzle-out (LONO) transfer (5-fold means). The MLP is consistently more conservative in coverage; the SVGP is the stronger point predictor. Physical baselines and per-fold detail are in Chapter~5.}
\label{tab:conclusions-scorecard}
\end{table}

\begin{itemize}
  \item On the conservative uncensored CDF protocol (the primary in-domain point-wise metric), the production three-stage MLP (anchor-off, \texttt{kd\_mse\_mu\_plus\_sigma} loss, \(\lambda_\sigma=5.0\), five seeds 7/17/42/99/2024) reduces RMSE by 59\% and 55\% relative to the calibrated Hiroyasu--Arai (\SI{10.192}{mm}) and delayed Naber--Siebers (\SI{9.212}{mm}) baselines refit on this same protocol (Chapter~5). A censoring-inclusive supplementary evaluation on the full cleaned set, which additionally includes the right-censored points, is reported in Chapter~5.
  \item In-distribution, the SVGP is the stronger point-wise regressor, but the MLP is markedly more conservative in uncertainty coverage (1$\sigma$/2$\sigma$ \(0.888/0.990\) versus \(0.717/0.946\)); the structural value of the MLP relative to the SVGP therefore lies in its conservative coverage rather than point accuracy.
  \item Under leave-one-nozzle-out transfer the SVGP wins the point RMSE in four of five folds; after excluding the off-design Nozzle0 fold where both architectures collapse (MLP RMSE \(\sim\SI{35}{mm}\), SVGP \(\sim\SI{27}{mm}\)), the 1$\sigma$ coverage gap closes (\(0.696\) versus \(0.697\)). Once the off-design Nozzle0 fold is excluded, both surrogates fall well below the legacy LONO-scale physical-baseline reference (MLP \(\SI{7.00}{mm}\), SVGP \(\SI{5.95}{mm}\), versus Hiroyasu--Arai \(\SI{10.713}{mm}\pm\SI{1.542}{mm}\) and Naber--Siebers \(\SI{9.784}{mm}\pm\SI{2.102}{mm}\); Chapter~5). Nozzle0 is the single out-of-design-family fold (a different injector hardware family) and it alone lifts the five-fold mean: with it included, the MLP (\(\SI{12.55}{mm}\)) sits above both baselines and the SVGP (\(\SI{10.16}{mm}\)) falls below Hiroyasu--Arai but not Naber--Siebers. This split is the inductive-bias behaviour discussed in Section~\ref{sec:kernel-vs-nn-ood-en}; the low coverage on difficult nozzles shows that extrapolation risk to genuinely off-design geometries must still be reported explicitly.
\end{itemize}

\subsection{Current Prototype Boundaries}

\begin{itemize}
  \item The probabilistic wall-interaction screening demonstrates that the penetration surrogate can already be projected into time-resolved geometric risk scores; however, the current rotated Gaussian geometry and toy piston case are merely pipeline demonstrations and have yet to be validated against matching experiments or CFD.
  \item The model should be understood as an early-stage screening surrogate within the non-reacting, room-temperature, Mie-defined operating envelope, rather than a phase-resolved predictor for liquid impingement in a real engine.
\end{itemize}

\subsection{Next Development Priorities}

\begin{enumerate}
  \item Validate $P_{\mathrm{hit}}(t)$ and cumulative exposure $E$ using matching optical experiments or CFD impingement data.
  \item Replace the current analytical constant-cone-angle envelope with a learnable two-dimensional envelope model.
  \item Report condition-level confidence intervals (clustered by recording, condition, and nozzle) and bootstrap confidence bands, since the 540,367 point-wise samples are temporally correlated within a trajectory and cannot be treated as independent.
  \item Introduce nozzle-conditioned transient priors so that difficult in-family cases such as Nozzle~3 (wake-driven slug/step-wise penetration) and the out-of-design-family Nozzle0 can be modeled more explicitly.
\end{enumerate}

\subsection{Closing Perspective}

The nozzle geometries and the two-millisecond Mie archive are specific to this application; the engineering sequence that turned them into a calibrated screening surrogate is not. Data extraction from a raw archive, censoring-aware curation, physics-informed target construction, staged training with knowledge distillation, held-out evaluation, and screening deployment are the same stages now documented for large-scale machine-learning systems~\cite{Zhao2023LLMSurvey}. Carrying that discipline end to end (at the modest scale of one researcher and one workstation) produced a conservative, uncertainty-aware surrogate that outperforms calibrated physical baselines on the primary uncensored-CDF metric by more than half, together with a probabilistic wall-impingement screener built on top of it. The scale and the modality differ from large-model engineering; the workflow does not.

\clearpage
