\documentclass[english, 12pt, a4paper, elec, utf8, a-2b, online]{aaltothesis}

\setupthesisfonts

\usepackage{graphicx}
\graphicspath{{images/}{../images/}{latex/../images/}}
\usepackage{array}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage[type={CC}, modifier={by-nc-sa}, version={4.0}]{doclicense}
\usepackage[style=ieee,sorting=none]{biblatex}
\IfFileExists{thesisreferences.bib}{%
  \addbibresource{thesisreferences.bib}%
}{%
  \IfFileExists{../thesisreferences.bib}{%
    \addbibresource{../thesisreferences.bib}%
  }{%
    \addbibresource{latex/../thesisreferences.bib}%
  }%
}

\urlstyle{tt}
\DeclareRobustCommand{\codepath}[1]{\nolinkurl{#1}}
\pdfstringdefDisableCommands{\def\codepath#1{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\setlength{\headheight}{36pt}
\setlength{\emergencystretch}{2em}
\newcommand{\thesisinput}[1]{%
  \IfFileExists{#1.tex}{%
    \input{#1}%
  }{%
    \IfFileExists{latex/#1.tex}{%
      \input{latex/#1}%
    }{%
      \input{../latex/#1}%
    }%
  }%
}

\newcommand{\setmainabstract}[1]{%
  \def\mainabstract{#1}%
  \thesisabstract{#1}%
}
\setmainabstract{Pilot diesel injections in multi-fuel marine engines are short, strongly transient events. If the liquid envelope reaches the piston crown or cylinder wall before useful ignition develops, wall wetting, lubricant dilution, and poor mixture preparation can follow. This thesis develops an evidence-grounded screening workflow that converts high-speed Mie-scattering videos into penetration-oriented observables and a surrogate model for early spray-wall-risk assessment. The data consist of room-temperature, non-reacting top-view recordings from the W\"artsil\"a Optical Spray Combustion Chamber, covering multiple operating conditions and nozzle geometries. The workflow combines manual nozzle calibration, CUDA-accelerated multihole video processing, plume-wise feature extraction, and a reduced-order trajectory fit that both smooths noisy labels and reduces the downward bias caused by right-censored observations near the field-of-view limit. On these fitted targets, a three-stage neural surrogate is trained. Stage 1 learns a deterministic representative-trajectory baseline. Stage 2 extends the model to heteroscedastic mean-and-variance prediction on the full filtered dataset. Stage 3 refines early-time behaviour against raw CDF observations while retaining conservative teacher guidance where late-time coverage becomes sparse, so that the model does not mistake disappearance from the image window for physical saturation. On 692,942 held-out prediction points from 24,316 trajectories across 9 nozzle campaigns under the full-clean CDF evaluation protocol, the final anchor-off Stage 3 model (hybrid mean-MSE plus log-variance-MSE distillation, sigma-weight 5.0) achieves a five-seed mean RMSE of 4.74 mm and MAE of 3.29 mm with a seed-to-seed standard deviation below 0.04 mm, while retaining 86.7 percent at 1 sigma and 98.5 percent at 2 sigma uncertainty coverage, and reduces RMSE by more than 55 percent against the calibrated Hiroyasu--Arai and delayed Naber--Siebers empirical baselines. A sparse heteroscedastic Gaussian-process (SVGP) is also evaluated under the same protocol as an alternative backbone: SVGP is the stronger point regressor both in-distribution (RMSE 4.63 mm) and under leave-one-nozzle-out transfer (5.95 mm versus the MLP's 7.00 mm on the four modern-design folds). The MLP backbone is retained for its more conservative uncertainty coverage, its onset side output used by the impingement-screening interface, and the ablation affordability that supports the design-attribution evidence at a master's-level compute budget; the two should be read as different accuracy--coverage trade-offs within the same censoring-aware learning protocol rather than as an architecture-level win. A simplified geometric projection then maps the predicted penetration distribution onto a probabilistic collision-screening proxy. The intended use is explicit: the workflow is an early-stage engineering screening tool within its measured operating domain, not a substitute for CFD, engine testing, or validated in-cylinder impingement physics.}

\degreeprogram{Electronics and Nanotechnology}
\major{Microelectronic Circuit Design}
\univdegree{MSc}
\thesisauthor{Linan Jiang}
\thesistitle{Data-Driven Penetration Prediction and Probabilistic Impingement Screening Using Mie Scattering Imaging}
\thesissubtitle{An uncertainty-aware surrogate workflow for multi-hole diesel sprays}
\place{Espoo}
\date{19 May 2026}
\supervisor{Prof.\ Ossi Kaario}
\advisor{Dr Zeeshan Ahmad}
\advisor{Dr Qiang Cheng}
\collaborativepartner{Wärtsilä}
\uselogo{'}
\copyrighttext{\noexpand\textcopyright\ \number\year. This work is
    licensed under a Creative Commons "Attribution-NonCommercial-ShareAlike 4.0
    International" (BY-NC-SA 4.0) license.}{\noindent\textcopyright\ \number
    \year\ \doclicenseThis}
\keywords{Mie scattering\spc spray imaging\spc penetration\spc computer vision\spc surrogate modeling}

\begin{document}

\makecoverpage
\makecopyrightpage

\begin{abstractpage}[english]
\mainabstract

\end{abstractpage}

\mysection{Preface}
Special thanks to Professor Ossi Kaario for supervision, Dr Zeeshan Ahmad, Mr Viljam Grahn,
and Dr Qiang Cheng for technical guidance, 
and all Wärtsilä collaborators around the spray-imaging workflow for helping shape the practical direction of this work.

\mysection{Use of AI Tools}
This thesis was prepared with assistance from AI-based tools for exploratory code implementation, script automation, data-screening support, summarising code and processing workflows, visualisation preparation, figure-preparation support, paraphrasing, grammar correction, and language polishing. AI tools were used within a layered and auditable research workflow: code-level outputs were checked through executable programs, logs, and intermediate files; data and figure-level outputs were cross-checked against exported tables, diagnostic plots, and numerical results; and text-level revisions were grounded in the preceding code, data, figures, literature, and author interpretation.

AI tools were not used as sources of scientific evidence, citation sources, sources of experimental results, or independent authors. The research questions, methodological choices, implementation decisions, interpretation of results, and final thesis text were reviewed, edited, and confirmed by the author. The author remains fully responsible for the accuracy, originality, code, analyses, figures, results, and conclusions presented in this thesis.

\thesistableofcontents
\dothesispagenumbering{}

\thesisinput{sections_en/01_introduction}
\thesisinput{sections_en/02_literature_review}
\thesisinput{sections_en/03_data_workflow}
\thesisinput{sections_en/04_trajectory_surrogate_screening}
\thesisinput{sections_en/05_results}
\thesisinput{sections_en/06_discussion}
\thesisinput{sections_en/07_conclusions}
\thesisinput{sections_en/appendix_implementation_mapping}
\thesisbibliography
\printbibliography

\end{document}
